Deriving Multi-Headed Planar Dependency Parses from Link Grammar Parses

### 2014-12-07 doc/slides/hong+eisner.TLT13.slides.tex
updated ILP slides

Variables:

\begin{itemize}
\item $x_{ij}, x_{ji} \in \mathbb{Z} \geq 0$ representing the orientation of each link

$x_{ij} + x_{ji} = 1$

@@

\subsection{Connectedness, Acyclicity}
\frame{\frametitle{Connectedness, Acyclicity}
\begin{itemize}
  \item[] Connectedness
    \begin{align}
      \sum_u x_{uv} & \geq 1
    \end{align}

  \item[] Acyclicity
    \begin{itemize}
      \item a
    \end{itemize}

@@

\frame{\frametitle{Consistence of Directionalized Links}
\begin{itemize}
  \item[] Consistency of Directionalized Links
    \begin{align}\label{direction+slack}
      x_{ij} &\leq r_L &
      x_{ji} &\leq \ell_L 
    \end{align}

    \begin{align}\label{eqn:obj}
      \min \left( \sum_L r_L + \ell_L \right)
    \end{align}

    %$$x_{ij} \leq r_L + s_{ij} $$
    %$$x_{ji} \leq \ell_L + s_{ij}$$
    %$$\min \left( \sum_L r_L + \ell_L \right)$$

@@

\frame{\frametitle{Consistence of Directionalized Links with Slack}
\begin{itemize}
  \item[] Consistency of Directionalized Links
    \begin{align}\label{direction+slack}
      x_{ij} &\leq r_L + s_{ij} &
      x_{ji} &\leq \ell_L + s_{ij} 
    \end{align}

    \begin{align}\label{eqn:obj}
      \min \left( \sum_L r_L + \ell_L \right) \cdot \frac{N_L}{4} + \sum_{ij}s_{ij}
    \end{align}

    %$$x_{ij} \leq r_L + s_{ij} $$
    %$$x_{ji} \leq \ell_L + s_{ij}$$
    %$$\min \left( \sum_L r_L + \ell_L \right) \cdot \frac{N_L}{4} + \sum_{ij}s_{ij}$$

@@

\frame{\frametitle{ILP Results}
  \begin{itemize}
    \item[] Link types that allowed both directions:
      \begin{itemize}
        \item[] $\frac{7}{113} = 6.19\%$ 
      \end{itemize}

    \item[] Link tokens that required disallowed direction via slack:
      \begin{itemize}
        \item[] $\frac{4043}{195000} = 2.07\%$ 
      \end{itemize}
  \end{itemize}

  \begin{itemize}
  \item[]
  \end{itemize}
}

\frame{\frametitle{ILP Results}
  \begin{itemize}
    \item[] Link types that allowed both directions:
      \begin{itemize}
        \item[] $\frac{7}{113} = 6.19\%$ 
      \end{itemize}

    \item[] Link tokens that required disallowed direction via slack:
      \begin{itemize}
        \item[] $\frac{4043}{195000} = 2.07\%$ 
      \end{itemize}
  \end{itemize}

  \begin{itemize}
  \item[] The link labels mostly have a consistent direction.
  \end{itemize}
}

\frame{\frametitle{Link Results}
  \begin{itemize}
    \item Subject-Verb links are backwards.
  \end{itemize}

\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	\end{deptext}
	\deproot[edge above, edge style = {blue, dotted}]{3}{\small ROOT}
	\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
	\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
	\depedge[edge above, edge style = {blue, dotted}]{3}{4}{\small ADV}
	\depedge[edge above, edge style = {blue, dotted}]{3}{5}{\small ADV}
	\depedge[edge below, edge style = {red, thick}]{3}{6}{I}
	\depedge[edge above, edge style = {blue, thick}]{3}{6}{\small VC}
	\depedge[edge above, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
	\depedge[edge below, edge style = {red, thick}]{2}{1}{D}
	\depedge[edge above, edge style = {blue, thick}]{2}{1}{\small NMOD}
	\depedge[edge below, edge style = {red, thick}]{7}{8}{MV}
	\depedge[edge above, edge style = {blue, thick}]{7}{8}{\small ADV}
	\depedge[edge below, edge style = {orange, thick}]{6}{7}{P}
	\depedge[edge above, edge style = {blue, thick}]{6}{7}{\small VC}
	\depedge[edge below, edge style = {red, thick}]{8}{9}{J}
	\depedge[edge above, edge style = {blue, thick}]{8}{9}{\small PMOD}
	\deproot[edge below, edge style = {red, dotted}]{2}{W}
	\deproot[edge below, edge style = {orange, ultra thick, dotted}]{7}{WV}
	\deproot[edge below, edge style = {red, dotted}]{10}{X}
	\depedge[edge below, edge style = {red, dotted}]{6}{4}{E}
	\depedge[edge below, edge style = {red, dotted}]{6}{5}{E}
  \end{dependency}
\end{figure}

@@

\frame{\frametitle{Link Results}
  \begin{itemize}
    \item Subject-Verb links are backwards.
  \end{itemize}

\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	\end{deptext}
	%\deproot[edge above, edge style = {blue, dotted}]{3}{\small ROOT}
	\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
	\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
	%\depedge[edge above, edge style = {blue, dotted}]{3}{4}{\small ADV}
	%\depedge[edge above, edge style = {blue, dotted}]{3}{5}{\small ADV}
	%\depedge[edge below, edge style = {red, thick}]{3}{6}{I}
	%\depedge[edge above, edge style = {blue, thick}]{3}{6}{\small VC}
	%\depedge[edge above, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
	%\depedge[edge below, edge style = {red, thick}]{2}{1}{D}
	%\depedge[edge above, edge style = {blue, thick}]{2}{1}{\small NMOD}
	%\depedge[edge below, edge style = {red, thick}]{7}{8}{MV}
	%\depedge[edge above, edge style = {blue, thick}]{7}{8}{\small ADV}
	%\depedge[edge below, edge style = {orange, thick}]{6}{7}{P}
	%\depedge[edge above, edge style = {blue, thick}]{6}{7}{\small VC}
	%\depedge[edge below, edge style = {red, thick}]{8}{9}{J}
	%\depedge[edge above, edge style = {blue, thick}]{8}{9}{\small PMOD}
	%\deproot[edge below, edge style = {red, dotted}]{2}{W}
	%\deproot[edge below, edge style = {orange, ultra thick, dotted}]{7}{WV}
	%\deproot[edge below, edge style = {red, dotted}]{10}{X}
	%\depedge[edge below, edge style = {red, dotted}]{6}{4}{E}
	%\depedge[edge below, edge style = {red, dotted}]{6}{5}{E}
  \end{dependency}
\end{figure}

>>>>

Variables: 
\begin{itemize}
\item[] $x_{ij}, x_{ji} \in \mathbb{Z} \geq 0$: orientation of each link
\item[] $x_{ij} + x_{ji} = 1$

@@

\frame{\frametitle{Connectedness, Acyclicity}
\begin{itemize}
  \item[] Connectedness
    \begin{align}
      \sum_u x_{uv} & \geq 1
    \end{align}

  \item[] Acyclicity
    \begin{itemize}
      \item[] Given that node $u$ is the parent of $v$
      \item[] $n_v$: length of the sentence containing node $v$
      \item[] $d_v \in [0, n_v]$: depth of the node from the root of the sentence
    \end{itemize}

@@

\frame{\frametitle{Consistency of Directionalized Links}
\begin{itemize}
  \item[] Consistency of Directionalized Links
    \begin{itemize}
      \item[] $r_L, \ell_L \in \{0,1\}$: whether links with label $L$ allowed left/right
    \end{itemize}
    \begin{align}\label{direction+slack}
      x_{ij} &\leq r_L &
      x_{ji} &\leq \ell_L 
    \end{align}
    \begin{align}\label{eqn:obj}
      \min \left( \sum_L r_L + \ell_L \right)
    \end{align}
    \begin{itemize}
      \item[]  
      \item[]  
    \end{itemize}

    \begin{itemize}
      \item[]
    \end{itemize}

@@

\frame{\frametitle{Consistency of Directionalized Links with Slack}
\begin{itemize}
  \item[] Consistency of Directionalized Links
    \begin{itemize}
      \item[] $r_L, \ell_L \in \{0,1\}$: whether links with label $L$ allowed left/right
    \end{itemize}
    \begin{align}\label{direction+slack}
      x_{ij} &\leq r_L + s_{ij} &
      x_{ji} &\leq \ell_L + s_{ij} 
    \end{align}
    \begin{align}\label{eqn:obj}
      \min \left( \sum_L r_L + \ell_L \right) \cdot \frac{N_L}{4} + \sum_{ij}s_{ij}
    \end{align}
    \begin{itemize}
      \item[] $s_{ij} \in \mathbb{R} \geq 0$: slack variable 
      \item[] $N_L$: Number of link tokens with label $L$
    \end{itemize}

    \begin{itemize}
      \item[] Slack allows a few links with label $L$ in disallowed directions
    \end{itemize}

@@

\frame{\frametitle{ILP Results}
  \begin{itemize}
    \item[] Link types that allowed both directions:
      \begin{itemize}
        \item[] $\frac{7}{113} = 6.19\%$ 
      \end{itemize}
    \item[] Link tokens that required disallowed direction via slack:
      \begin{itemize}
        \item[] $\frac{4043}{195000} = 2.07\%$ 
      \end{itemize}
  \end{itemize}
  \begin{itemize}
  \item[]  
  \end{itemize}
}

\frame{\frametitle{ILP Results}
  \begin{itemize}
    \item[] Link types that allowed both directions:
      \begin{itemize}
        \item[] $\frac{7}{113} = 6.19\%$ 
      \end{itemize}
    \item[] Link tokens that required disallowed direction via slack:
      \begin{itemize}
        \item[] $\frac{4043}{195000} = 2.07\%$ 
      \end{itemize}
  \end{itemize}
  \begin{itemize}
  \item[] The link labels mostly have a consistent direction.
  \end{itemize}
}

\frame{\frametitle{Link Results: Subject-Verb links are backwards}
\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	\end{deptext}
	\deproot[edge above, edge style = {blue, dotted}]{3}{\small ROOT}
	\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
	\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
	\depedge[edge above, edge style = {blue, dotted}]{3}{4}{\small ADV}
	\depedge[edge above, edge style = {blue, dotted}]{3}{5}{\small ADV}
	\depedge[edge below, edge style = {red, thick}]{3}{6}{I}
	\depedge[edge above, edge style = {blue, thick}]{3}{6}{\small VC}
	\depedge[edge above, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
	\depedge[edge below, edge style = {red, thick}]{2}{1}{D}
	\depedge[edge above, edge style = {blue, thick}]{2}{1}{\small NMOD}
	\depedge[edge below, edge style = {red, thick}]{7}{8}{MV}
	\depedge[edge above, edge style = {blue, thick}]{7}{8}{\small ADV}
	\depedge[edge below, edge style = {orange, thick}]{6}{7}{P}
	\depedge[edge above, edge style = {blue, thick}]{6}{7}{\small VC}
	\depedge[edge below, edge style = {red, thick}]{8}{9}{J}
	\depedge[edge above, edge style = {blue, thick}]{8}{9}{\small PMOD}
	\deproot[edge below, edge style = {red, dotted}]{2}{W}
	\deproot[edge below, edge style = {orange, ultra thick, dotted}]{7}{WV}
	\deproot[edge below, edge style = {red, dotted}]{10}{X}
	\depedge[edge below, edge style = {red, dotted}]{6}{4}{E}
	\depedge[edge below, edge style = {red, dotted}]{6}{5}{E}
  \end{dependency}
\end{figure}

@@

\frame{\frametitle{Link Results: Subject-Verb links are backwards}
\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	\end{deptext}
	\deproot[edge above]{3}{\small ROOT}
	\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
	\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
	\depedge[edge above]{3}{4}{\small ADV}
	\depedge[edge above]{3}{5}{\small ADV}
	\depedge[edge below]{3}{6}{I}
	\depedge[edge above]{3}{6}{\small VC}
	\depedge[edge above, edge unit distance =1.5ex]{3}{10}{\small P}
	\depedge[edge below]{2}{1}{D}
	\depedge[edge above]{2}{1}{\small NMOD}
	\depedge[edge below]{7}{8}{MV}
    \depedge[edge above]{7}{8}{\small ADV}
	\depedge[edge below]{6}{7}{P}
	\depedge[edge above]{6}{7}{\small VC}
	\depedge[edge below]{8}{9}{J}
	\depedge[edge above]{8}{9}{\small PMOD}
	\deproot[edge below]{2}{W}
    \deproot[edge below]{7}{WV}
	\deproot[edge below]{10}{X}
	\depedge[edge below]{6}{4}{E}
	\depedge[edge below]{6}{5}{E}
  \end{dependency}
\end{figure}
